### 2024-01-20 outputs/paper/paper.qmd
Removed the llm folder and updated files in the paper folder

---
title: "My title"
author: 
  - Benny Rochwerg
thanks: "A GitHub Repository containing all data, R code, and other files used in this investigation is located here: <https://github.com/bennyrochwerg/healthcare-outbreaks-toronto>"
date: 01-23-2024
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

# Data

## Sketches

Sketches depicting both the desired dataset and some of the graphs that will be generated in this analysis are available in the GitHub Repository.

The R programming language was used to conduct this analysis [@citeR], along with the `tidyverse` package [@citetidyverse].

# Acknowledgements

Some of the R code used to create this work was adapted from @tellingstorieswithdata.

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("outputs/models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-01-20 outputs/paper/paper.qmd
Downloaded and saved the raw dataset in the inputs folder

@@
 
-The R programming language was used to conduct this analysis [@citeR], along with the `tidyverse` package [@citetidyverse].
+The R programming language was used to conduct this analysis [@citeR], along with the `janitor` [@citejanitor], `tidyverse` [@citetidyverse], and `opendatatoronto` [@citeopendatatoronto] packages.
 

### 2024-01-20 scripts/01-download_data.R

@@ -1,26 +1,25 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Downloads the data from the opendatatoronto package (Gelfand 2022)
+# and saves the dataset for use in the main paper.
+# Author: Benny Rochwerg
+# Date: January 23, 2024
+# Contact: 4321benny@gmail.com
+# Pre-requisites: Installation of the opendatatoronto (Gelfand 2022) and
+# tidyverse (Wickham et al. 2019) packages.
 
+#### Loading Packages ####
 
-#### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
 
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+#### Downloading the Dataset ####
 
+outbreak_raw_data <-
+  search_packages("Outbreaks in Toronto Healthcare Institutions") |>
+  list_package_resources() |>
+  filter(name == "ob_report_2023") |>
+  get_resource()
 
+#### Saving the Dataset ####
 
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
-
-         
+write_csv(outbreak_raw_data, "inputs/data/raw_data.csv")

### 2024-01-20 scripts/02-data_cleaning.R
Cleaned the raw dataset and saved the cleaned dataset as a CSV file

@@ -1,44 +1,47 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Cleans the raw Toronto healthcare outbreaks dataset obtained in
+# "01-download_data.R".
+# Author: Benny Rochwerg
+# Date: January 23, 2024
+# Contact: 4321benny@gmail.com
+# Pre-requisites: Run the file "01-download_data.R".
 
-#### Workspace setup ####
+#### Loading Packages ####
+
 library(tidyverse)
 
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+#### Cleaning the Dataset ####
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+outbreak_raw_data <- read_csv("inputs/data/raw_data.csv")
 
-#### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+# This code is based on knowledge from Alexander (2023).
+
[…]

### 2024-01-20 scripts/02-data_cleaning.R
Cleaned the Outbreak First Known Case column

@@ -29,19 +29,39 @@ outbreak_cleaned_data <-
          `Outbreak First Known Cause` = `Causative Agent-1`) |>
   
   # Renaming the Type of Location entries
-  mutate(`Type of Location` = case_match(`Type of Location`,
-                                         "LTCH" ~
-                                           "Long-Term Care Home",
-                                         "Hospital-Acute Care" ~
-                                           "Hospital (Acute Care)",
-                                         "Retirement Home" ~
-                                           "Retirement Home",
-                                         "Hospital-Chronic Care" ~
-                                           "Hospital (Chronic Care)",
-                                         "Hospital-Psychiatric" ~
-                                           "Hospital (Psychiatric)",
-                                         "Transitional Care" ~
-                                           "Transitional Care"))
+  mutate(`Type of Location` =
+           case_match(`Type of Location`,
+                      "LTCH" ~ "Long-Term Care Home",
+                      "Hospital-Acute Care" ~ "Hospital (Acute Care)",
+                      "Retirement Home" ~ "Retirement Home",
+                      "Hospital-Chronic Care" ~ "Hospital (Chronic Care)",
+                      "Hospital-Psychiatric" ~ "Hospital (Psychiatric)",
+                      "Transitional Care" ~ "Transitional Care")) |>
+  
+  # Renaming some of the Outbreak First Known Cause entries for simplicity
+  mutate(`Outbreak First Known Cause` =
+           case_match(`Outbreak First Known Cause`,
+                      "COVID-19" ~ "COVID-19",
+                      "Parainfluenza" ~ "Parainfluenza",
+                      "Respiratory syncytial virus" ~
+                        "Respiratory syncytial virus",
+                      "Metapneumovirus" ~ "Metapneumovirus",
+                      "Norovirus" ~ "Norovirus",
+                      "Rhinovirus" ~ "Rhinovirus",
+                      "Group B Streptococcal disease (neonatal)" ~
+                        "Group B Streptococcal disease (neonatal)",
+                      "Enterovirus/Rhinovirus" ~ "Enterovirus/Rhinovirus",
+                      "Enterovirus" ~ "Enterovirus",
+                      "Streptococcus pyogenes" ~ "Streptococcus pyogenes",
+                      "Parainfluenza PIV III" ~ "Parainfluenza",
[…]

### 2024-01-20 scripts/02-data_cleaning.R
Added a blank line for formatting purposes

@@ -64,4 +64,5 @@ outbreak_cleaned_data <-
                       "Unable to identify" ~ "Pending/Unknown"))
 
 #### Saving the Cleaned Dataset ####
+
 write_csv(outbreak_cleaned_data, "outputs/data/cleaned_data.csv")

### 2024-01-20 scripts/00-simulate_data.R
Ran tests on the cleaned dataset

@@ -5,7 +5,7 @@
 # Author: Benny Rochwerg
 # Date: January 23, 2024
 # Contact: 4321benny@gmail.com
-# Pre-requisites: Installation of the janitor (Firke 2023) and tidyverse
+# Pre-requisites: Install the janitor (Firke 2023) and tidyverse
 # (Wickham et al. 2019) packages.
 
 #### Loading Packages ####
@@ -33,7 +33,7 @@ simulated_outbreak_data <-
     
     # Choosing a type of outbreak 1000 times at random with replacement
     "Type of Outbreak" = sample(
-      x = c("Respiratory", "Enteric"),
+      x = c("Respiratory", "Enteric", "Other"),
       size = 1000,
       replace = TRUE),
     
@@ -49,8 +49,7 @@ simulated_outbreak_data <-
 
 # This code was adapted from Alexander (2023).
 
-# Checking the class of each column
-simulated_outbreak_data$ID |> class() == "integer"
+# Checking the class of some of the columns
 
 simulated_outbreak_data$`Type of Location` |> class() == "character"
 
@@ -61,15 +60,16 @@ simulated_outbreak_data$`Outbreak First Known Cause` |> class() == "character"
 # Checking that the minimum value in the ID column is at least 1
 simulated_outbreak_data$ID |> min() >= 1
 
-# Checking that the only two outbreak types present are Respiratory and Enteric
+# Checking that the only three outbreak types present are Respiratory, Enteric,
+# and Other
 simulated_outbreak_data$`Type of Outbreak` |>
   unique() |>
-  sort() == sort(c("Respiratory", "Enteric"))
+  sort() == sort(c("Respiratory", "Enteric", "Other"))
 
-# Checking that only two outbreak types are present
+# Checking that only three outbreak types are present
 simulated_outbreak_data$`Type of Outbreak` |>
   unique() |>
-  length() == 2
+  length() == 3
 
 # Checking that the only four location types present are Long-Term Care Home,
 # Retirement Home, Hospital (Psychiatric), and Other

### 2024-01-20 scripts/01-download_data.R

@@ -4,7 +4,7 @@
 # Author: Benny Rochwerg
 # Date: January 23, 2024
 # Contact: 4321benny@gmail.com
-# Pre-requisites: Installation of the opendatatoronto (Gelfand 2022) and
+# Pre-requisites: Install the opendatatoronto (Gelfand 2022) and
 # tidyverse (Wickham et al. 2019) packages.
 
 #### Loading Packages ####

### 2024-01-20 scripts/02-data_cleaning.R

@@ -61,7 +61,10 @@ outbreak_cleaned_data <-
                       "CPE Enterobacter unspecified (NDM)" ~ "CPE",
                       "CPE Unspecified (KPC)" ~ "CPE",
                       "Pending" ~ "Pending/Unknown",
-                      "Unable to identify" ~ "Pending/Unknown"))
+                      "Unable to identify" ~ "Pending/Unknown")) |>
+  
+  # Changing the class of the ID column to "integer"
+  mutate(`ID` = as.integer(`ID`))
 
 #### Saving the Cleaned Dataset ####
 

### 2024-01-20 scripts/03-test_data.R

@@ -1,15 +1,44 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Runs tests on the cleaned dataset to check the validity
+# of the entries.
+# Author: Benny Rochwerg
+# Date: January 23, 2024
+# Contact: 4321benny@gmail.com
+# Pre-requisites: Run the files "01-download_data.R" and "02-data_cleaning.R".
 
+#### Loading Packages ####
 
-#### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
-#### Test data ####
+#### Testing the Cleaned Dataset ####
+
+# This code was adapted from Alexander (2023).
+# Note that only the tests from "00-simulate_data.R" that are applicable to this
+# cleaned dataset will be performed.
+
+# Loading the cleaned dataset
+outbreak_cleaned_data <- read_csv("outputs/data/cleaned_data.csv")
+
+# Running the tests on the cleaned dataset
+
+# Checking the class of some of the columns
+
+outbreak_cleaned_data$`Type of Location` |> class() == "character"
+
+outbreak_cleaned_data$`Type of Outbreak` |> class() == "character"
+
+outbreak_cleaned_data$`Outbreak First Known Cause` |> class() == "character"
+
+# Checking that the minimum value in the ID column is at least 1
+outbreak_cleaned_data$ID |> min() >= 1
+
+# Checking that the only three outbreak types present are Respiratory, Enteric,
+# and Other
+outbreak_cleaned_data$`Type of Outbreak` |>
+  unique() |>
+  sort() == sort(c("Respiratory", "Enteric", "Other"))
+
+# Checking that only three outbreak types are present
+outbreak_cleaned_data$`Type of Outbreak` |>
+  unique() |>
+  length() == 3

### 2024-01-21 outputs/paper/paper.qmd
Loaded the tidyverse package in the paper.qmd file

@@
 
-# Introduction
+```{r}
+#### Loading Packages ####
 
-# Data
+# install.packages("tidyverse")
+library(tidyverse)
+```
+
+# Introduction {#introduction}
+
+# Data {#data}
 
@@
 Sketches depicting both the desired dataset and some of the graphs that will be generated in this analysis are available in the GitHub Repository.
+
+## Visualizing the data
+
+```{r}
+#### Creating Tables and Graphs ####
+
+# Loading the cleaned dataset
+
+# outbreak_cleaned_data <- read_csv("outputs/data/cleaned_data.csv")
+
+# Creating a graph to depict the number of outbreaks by location type
+
+#| label: num-outbreaks-by-location
+#| fig-cap: Number of Toronto healthcare outbreaks by type of location in 2023
+#| echo: false
+
+# Obtaining the number of outbreaks by type of location
+# The knowledge used to write this code was gained in STA304H1.
+
+```
 

### 2024-01-21 outputs/paper/paper.qmd
Updated the Data section of the paper

--- 2024-01-20 outputs/paper/paper.qmd
@@
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 
-# Introduction
+```{r}
+#| echo: false
+#| warning: false
+#### Loading Packages ####
 
-# Data
+# install.packages("tidyverse")
+# install.packages("here")
+# install.packages("knitr")
+library(tidyverse)
+library(here)
+library(knitr)
+```
 
-## Sketches
+# Introduction {#sec-introduction}
 
-Sketches depicting both the desired dataset and some of the graphs that will be generated in this analysis are available in the GitHub Repository.
+# Data {#sec-data}
+
+## Overview {#sec-data-overview}
+
+The dataset used in this analysis is the 2023 installment of "Outbreaks in Toronto Healthcare Institutions" from Toronto Public Health (@tphdataset).
+In Ontario, the law mandates that respiratory and gastroenteric infections in healthcare facilities are followed and that any outbreaks, regardless of verification, are reported to public health entities such as Toronto Public Health (@tphdataset).
+Based on this information, Toronto Public Health updates this dataset on a weekly basis (@tphdataset).
+Moreover, this dataset is considered to be "open data" (@tphdataset) and can be utilized for a variety of purposes as long as an attribution statement (@sec-appendix-attribution) is attached and the link to the City of Toronto's Open Data License is added (@tphlicense).
+
+The variables included in this analysis are "Type of Location" (named "Outbreak Setting" in the original dataset), which refers to the nature of the healthcare offered at each facility; "Type of Outbreak", which indicates whether each outbreak is respiratory, enteric, or other in nature; and "Outbreak First Known Cause" (named "Causative Agent-1" in the original dataset), which denotes the first discovered pathogen based on at least one of the outbreak's cases (@tphdataset).
+
+Only two other datasets involving outbreaks, titled "COVID-19 Cases in Toronto" and "Daily Shelter & Overnight Service Occupancy & Capacity", could be identified following a search of the City of Toronto's Open Data Catalogue (@tphsearch).
+However, none of these datasets pertain specifically to healthcare settings in Toronto and were thus not used in this analysis.
+
+## Results {#sec-data-results}
+
+After loading the dataset using the R programming language [@citeR] and the `here` package [@citehere], (other packages), tables and graphs were created using R code adapted from @tellingstorieswithdata.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+#### Loading the Cleaned Dataset ####
+
+outbreak_cleaned_data <- read_csv(here::here("outputs/data/cleaned_data.csv"))
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: tbl-outbreaks-by-location-type
+#| tbl-cap: "Outbreaks by type of healthcare location in Toronto in 2023"
+#| tbl-colwidths: [50, 20, 20]
+
+#### First Table ####
+
+# This code is based on knowledge from Alexander (2023).
+
+outbreak_cleaned_data |>
+  count(`Type of Location`) |>
+  rename(`Type of location` = `Type of Location`,
+         `Number of outbreaks` = n) |>
+  mutate(`Proportion of total outbreaks` =
+           round(`Number of outbreaks` / nrow(outbreak_cleaned_data), 2)) |>
+  arrange(desc(`Number of outbreaks`)) |>
+  kable()
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: fig-outbreaks-by-location-type
+#| fig-cap: "Number of outbreaks by type of healthcare location in Toronto in 2023"
+
+#### First Graph ####
+
+# This code is based on knowledge from Alexander (2023).
+
+outbreak_cleaned_data |>
+  count(`Type of Location`) |>
+  # The reorder function used for sorting the columns in descending order was obtained
+  # from Stack Overflow:
+  # https://stackoverflow.com/questions/25664007/reorder-bars-in-geom-bar-ggplot2-by-value
+  ggplot(mapping = aes(x = reorder(`Type of Location`, n), y = n)) +
+  geom_col() +
+  theme_bw() +
+  labs(x = "Type of location", y = "Number of outbreaks") +
+  coord_flip()
+```
+
+@tbl-outbreaks-by-location-type and @fig-outbreaks-by-location-type illustrate that long-term care homes comprise a majority (approximately 56\%) of the total number of healthcare outbreaks in Toronto.
+This is followed by retirement homes (approximately 18\%), chronic care settings in hospitals (approximately 12\%), acute care settings in hospitals (approximately 10\%), psychiatric care settings in hospitals (approximately 2\%), and transitional care locations (approximately 1\%).
+Note that these percentages do not add up to 100\% as the proportions in @tbl-outbreaks-by-location-type were rounded to two decimal places.
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: tbl-outbreaks-by-outbreak-type
+#| tbl-cap: "Outbreaks by type of outbreak for Toronto healthcare locations in 2023"
+#| tbl-colwidths: [50, 20, 20]
+
+#### Second Table ####
+
+# This code is based on knowledge from Alexander (2023).
+
+outbreak_cleaned_data |>
+  count(`Type of Outbreak`) |>
+  rename(`Type of outbreak` = `Type of Outbreak`,
+         `Number of outbreaks` = n) |>
+  mutate(`Proportion of total outbreaks` =
+           round(`Number of outbreaks` / nrow(outbreak_cleaned_data), 2)) |>
+  arrange(desc(`Number of outbreaks`)) |>
+  kable()
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: fig-outbreaks-by-outbreak-type
+#| fig-cap: "Number of outbreaks by type of outbreak for Toronto healthcare locations in 2023"
+
+#### Second Graph ####
+
+# This code is based on knowledge from Alexander (2023).
+
+outbreak_cleaned_data |>
+  count(`Type of Outbreak`) |>
+  # The reorder function used for sorting the columns in descending order was obtained
+  # from Stack Overflow:
+  # https://stackoverflow.com/questions/25664007/reorder-bars-in-geom-bar-ggplot2-by-value
+  ggplot(mapping = aes(x = reorder(`Type of Outbreak`, -n), y = n)) +
+  geom_col() +
+  theme_bw() +
+  labs(x = "Type of outbreak", y = "Number of outbreaks")
+```
+
+@tbl-outbreaks-by-outbreak-type and @fig-outbreaks-by-outbreak-type show that the vast majority of healthcare outbreaks in Toronto are respiratory in nature (approximately 95\%).
+This is followed by enteric outbreaks (approximately 4\%) and other types of outbreaks (less than 1\%).
+Note that these percentages do not add up to 100\% as the proportions in @tbl-outbreaks-by-outbreak-type were rounded to two decimal places.
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: tbl-outbreaks-by-first-known-cause
+#| tbl-cap: "Outbreaks by first known cause for Toronto healthcare locations in 2023"
+#| tbl-colwidths: [50, 20, 20]
+
+#### Third Table ####
+
+# This code is based on knowledge from Alexander (2023).
+
+outbreak_cleaned_data |>
+  count(`Outbreak First Known Cause`) |>
+  rename(`Outbreak first known cause` = `Outbreak First Known Cause`,
+         `Number of outbreaks` = n) |>
+  mutate(`Proportion of total outbreaks` =
+           round(`Number of outbreaks` / nrow(outbreak_cleaned_data), 2)) |>
+  arrange(desc(`Number of outbreaks`)) |>
+  kable()
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: fig-outbreaks-by-first-known-cause
+#| fig-cap: "Number of outbreaks by first known cause for Toronto healthcare locations in 2023"
+
+#### Third Graph ####
+
+# This code is based on knowledge from Alexander (2023).
+
+outbreak_cleaned_data |>
+  count(`Outbreak First Known Cause`) |>
+  # The reorder function used for sorting the columns in descending order was obtained
+  # from Stack Overflow:
+  # https://stackoverflow.com/questions/25664007/reorder-bars-in-geom-bar-ggplot2-by-value
+  ggplot(mapping = aes(x = reorder(`Outbreak First Known Cause`, n), y = n)) +
+  geom_col() +
+  theme_bw() +
+  labs(x = "Outbreak first known cause", y = "Number of outbreaks") +
+  coord_flip()
+```
+
+@tbl-outbreaks-by-first-known-cause and @fig-outbreaks-by-first-known-cause illustrate that the majority of Toronto healthcare outbreaks have COVID-19 as their first known cause (approximately 66\%).
+This is followed by rhinovirus (approximately 7\%), respiratory syncytial virus (approximately 5\%), parainfluenza (approximately 3\%), influenza A (approximately 3\%), seasonal coronavirus (approximately 3\%), metapneumovirus (approximately 2\%), norovirus (approximately 2\%), enterovirus/rhinovirus (approximately 1\%), enterovirus (approximately 1\%), CPE (less than 1\%), neonatal Group B streptococcal disease (less than 1\%), and streptococcus pyogenes (less than 1\%).
+Approximately 7\% of outbreaks had a pending or unknown first known cause.
+Note that these percentages do not add up to 100\% as the proportions in @tbl-outbreaks-by-first-known-cause were rounded to two decimal places.
 
@@
 
-# Acknowledgements
+# Acknowledgements {#sec-acknowledgements}
 
 Some of the R code used to create this work was adapted from @tellingstorieswithdata.
-
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
-library(tidyverse)
-library(palmerpenguins)
-```
 
@@
 
-# Data {#sec-data}
+## Discussion {#data-discussion}
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+In @sec-data-results, it was observed that long-term care homes had the greatest proportion of total outbreaks among healthcare facilities in Toronto in 2023 (@tbl-outbreaks-by-location-type and @fig-outbreaks-by-location-type), that approximately 95\% of outbreaks were respiratory in nature (@tbl-outbreaks-by-outbreak-type and @fig-outbreaks-by-outbreak-type), and that COVID-19 was the first known cause of a majority of outbreaks (@tbl-outbreaks-by-first-known-cause and @fig-outbreaks-by-first-known-cause).
 
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
+All three of these results share common origins.
+For instance, in October 2022, the Government of Ontario removed a masking requirement for caregivers and visitors in long-term care homes [@cbcltch], whose residents are primarily at or above 65 years of age as reported in 2018 [@clri].
+Since individuals aged 60 years or older are more likely to experience worse health effects after contracting COVID-19, the Public Health Agency of Canada recommends that they wear a respirator or mask to limit the transmission of this pathogen [@phaccovid].
+As a result, it seems that the Government of Ontario did not heed this advice, possibly contributing to the large proportion of disease outbreaks (which include COVID-19) in Toronto long-term care homes in 2023.
+This appears to follow a long-standing trend in Ontario long-term care homes.
+For example, at the start of the COVID-19 pandemic, long-term care homes in Ontario experienced a sizable increase in COVID-19 cases [@cbcinquiry].
+This was potentially made worse by the fact that not all of these facilities enforced mask-wearing and that long-term care homes were passed over for personal protective equipment in favour of hospitals [@cbcinquiry].
+Moreover, the fact that several hospitals in Toronto curtailed their mask requirements in June and July of 2023 [@hospitalmasks] likely contributed to the large volume of respiratory disease outbreaks observed in @tbl-outbreaks-by-outbreak-type and @fig-outbreaks-by-outbreak-type.
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
+Although over 1,000 data points pertaining to healthcare outbreaks in Toronto in 2023 were examined in this analysis, several limitations were present.
+For instance, asymptomatic disease outbreaks were likely not included in the studied dataset as they may not have been noticeable enough to be reported, contributing to an underestimate of the total number of outbreaks.
+In addition, only the first known cause of each outbreak was examined for clarity purposes.
+This also likely led to an underestimate of the number of outbreaks associated with each disease.
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("outputs/models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+Future research should be conducted to assess the state of disease outbreaks in long-term care homes in particular, both in Toronto and across Canada, in order to gain deeper insight into this critical issue.
+The prevalence of disease outbreaks in other settings, such as schools and workplaces, should also be investigated to expand on this analysis.
 
@@
 
-# Appendix {-}
+# Appendix {#sec-appendix}
 
-# Additional data details
+## Dataset and Graph Sketches {#sec-appendix-sketches}
 
-# Model details {#sec-model-details}
+Sketches depicting both the desired dataset and the graphs to be generated in this analysis are available in the GitHub Repository.
 
-## Posterior predictive check
+## Attribution Statement {#sec-appendix-attribution}
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
+"Contains information licensed under the Open Government Licence – Toronto" (@tphlicense).
 
@@
 
-# References
+# References {#sec-references}

### 2024-01-21 scripts/02-data_cleaning.R

@@ -62,9 +62,6 @@ outbreak_cleaned_data <-
                       "CPE Unspecified (KPC)" ~ "CPE",
                       "Pending" ~ "Pending/Unknown",
                       "Unable to identify" ~ "Pending/Unknown")) |>
-  
-  # Changing the class of the ID column to "integer"
-  mutate(`ID` = as.integer(`ID`))
 
 #### Saving the Cleaned Dataset ####
 

### 2024-01-21 scripts/02-data_cleaning.R
Updated the data cleaning file to remove an unnecessary step

@@ -61,7 +61,7 @@ outbreak_cleaned_data <-
                       "CPE Enterobacter unspecified (NDM)" ~ "CPE",
                       "CPE Unspecified (KPC)" ~ "CPE",
                       "Pending" ~ "Pending/Unknown",
-                      "Unable to identify" ~ "Pending/Unknown")) |>
+                      "Unable to identify" ~ "Pending/Unknown"))
 
 #### Saving the Cleaned Dataset ####
 

### 2024-01-21 outputs/paper/paper.qmd
Finished writing the Data section

@@
 
-The dataset used in this analysis is the 2023 installment of "Outbreaks in Toronto Healthcare Institutions" from Toronto Public Health (@tphdataset).
-In Ontario, the law mandates that respiratory and gastroenteric infections in healthcare facilities are followed and that any outbreaks, regardless of verification, are reported to public health entities such as Toronto Public Health (@tphdataset).
-Based on this information, Toronto Public Health updates this dataset on a weekly basis (@tphdataset).
-Moreover, this dataset is considered to be "open data" (@tphdataset) and can be utilized for a variety of purposes as long as an attribution statement (@sec-appendix-attribution) is attached and the link to the City of Toronto's Open Data License is added (@tphlicense).
+The dataset used in this analysis is the 2023 installment of "Outbreaks in Toronto Healthcare Institutions" from Toronto Public Health [@tphdataset].
+In Ontario, the law mandates that respiratory and gastroenteric infections in healthcare facilities are followed and that any outbreaks, regardless of verification, are reported to public health entities such as Toronto Public Health [@tphdataset].
+Based on this information (or these measurements), Toronto Public Health updates this dataset on a weekly basis [@tphdataset].
+Moreover, this dataset is considered to be "open data" [@tphdataset] and can be utilized for a variety of purposes as long as an attribution statement is attached (@sec-appendix-attribution) and the link to the City of Toronto's Open Data License is added [@tphlicense].
 
-The variables included in this analysis are "Type of Location" (named "Outbreak Setting" in the original dataset), which refers to the nature of the healthcare offered at each facility; "Type of Outbreak", which indicates whether each outbreak is respiratory, enteric, or other in nature; and "Outbreak First Known Cause" (named "Causative Agent-1" in the original dataset), which denotes the first discovered pathogen based on at least one of the outbreak's cases (@tphdataset).
+The variables or measurements included in this analysis are "Type of Location" (named "Outbreak Setting" in the original dataset), which refers to the nature of the healthcare offered at each facility; "Type of Outbreak", which indicates whether each outbreak is respiratory, enteric, or other in nature; and "Outbreak First Known Cause" (named "Causative Agent-1" in the original dataset), which denotes the first discovered pathogen based on at least one of the outbreak's cases [@tphdataset].
 
-Only two other datasets involving outbreaks, titled "COVID-19 Cases in Toronto" and "Daily Shelter & Overnight Service Occupancy & Capacity", could be identified following a search of the City of Toronto's Open Data Catalogue (@tphsearch).
+Only two other datasets involving outbreaks, titled "COVID-19 Cases in Toronto" and "Daily Shelter & Overnight Service Occupancy & Capacity", could be identified following a search of the City of Toronto's Open Data Catalogue [@tphsearch].
 However, none of these datasets pertain specifically to healthcare settings in Toronto and were thus not used in this analysis.
 
+Using the R programming language [@citeR], the `janitor` [@citejanitor] and `tidyverse` [@citetidyverse] packages were used to simulate the dataset and generate tests for it.
+The `opendatatoronto` [@citeopendatatoronto] and `tidyverse` [@citetidyverse] packages were then applied in order to download the raw Toronto Public Health dataset.
+Next, the `tidyverse` package [@citetidyverse] was used to clean the raw dataset and test the cleaned dataset.
+
 ## Results {#sec-data-results}
-
-After loading the dataset using the R programming language [@citeR] and the `here` package [@citehere], (other packages), tables and graphs were created using R code adapted from @tellingstorieswithdata.
 
@@
 ```
+
+After loading the dataset using the R programming language [@citeR] and the `here` package [@citehere], the `tidyverse` [@citetidyverse] and `knitr` [@citeknitr] packages were used to generate tables and graphs.
+In doing so, R code was adapted from @tellingstorieswithdata.
 
@@
 
-The R programming language was used to conduct this analysis [@citeR], along with the `janitor` [@citejanitor], `tidyverse` [@citetidyverse], and `opendatatoronto` [@citeopendatatoronto] packages.
-
 # Acknowledgements {#sec-acknowledgements}
@@
 
-"Contains information licensed under the Open Government Licence – Toronto" (@tphlicense).
+"Contains information licensed under the Open Government Licence – Toronto" [@tphlicense].
 

### 2024-01-22 outputs/paper/paper.qmd
Finished the Abstract and Introduction sections of the paper

@@
 ---
-title: "My title"
+title: "Long-term care homes were hit hardest by 2023 disease outbreaks in Toronto healthcare facilities"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "In this report, Toronto Public Health data on 2023 outbreaks in healthcare facilities was analyzed.
+Long-term care homes comprised a majority of the outbreaks, approximately 95% of the outbreaks were respiratory, and almost two-thirds of the outbreaks had COVID-19 as their first known cause.
+This highlights the need for enhanced public health measures in healthcare facilities, particularly in long-term care homes.
+Future research should assess disease outbreaks in additional settings, such as workplaces and schools, to gain further insight into the extent of this issue."
 format: pdf
 number-sections: true
+# The toc parameter used to create a table of contents was obtained from
+# Quarto documentation:
+# https://quarto.org/docs/reference/formats/pdf.html
 toc: true
+# The tbl-colwidths parameter used to adjust the table column widths was
+# obtained from Quarto documentation:
+# https://quarto.org/docs/authoring/tables.html
+tbl-colwidths: [50, 20, 20]
 bibliography: references.bib
@@
 # Introduction {#sec-introduction}
+
+In March 2020, the World Health Organization proclaimed that the spread of COVID-19 constituted a pandemic [@whocovid].
+Shortly thereafter, in May 2020, the COVID-19 mortality rate in Canadian long-term care homes reached approximately 35\% and represented 81\% of Canadian COVID-19 deaths [@cihi].
+In general, the risk of disease circulation is greater in healthcare settings due to an increased presence of ill individuals and nearer physical contact between patients and staff [@cdc].
+
+Consequently, with the recent advent of large volumes of respiratory diseases such as influenza and COVID-19 in Toronto [@covidflu], assessing the prevalence of disease outbreaks in healthcare facilities in Toronto is of paramount importance.
+As a result, this investigation examines the number of outbreaks by location type (long-term care homes, retirement homes, hospitals, and transitional care facilities), outbreak type (respiratory and enteric/gastroenteric, the latter referring to intestinal and/or stomach illness [@enteric; @gastroenteric]), and first known cause (e.g., COVID-19, influenza A, etc.) to fill this knowledge gap.
+
+To do this, Toronto Public Health data on 2023 outbreaks in healthcare facilities was first obtained as described in @sec-data-overview.
+Based on this dataset, it was discovered that long-term care homes comprised a majority of the outbreaks, that approximately 95\% of the outbreaks were respiratory, and that almost two-thirds of the outbreaks had COVID-19 as their first known cause (@sec-data-results).
+As discussed in @sec-data-discussion, the high prevalence of outbreaks in long-term care homes has likely been exacerbated by the removal of a universal masking requirement, suggesting that the Government of Ontario should consider reinstating this policy to protect these vulnerable individuals.
+With respect to the structure of this paper, @sec-data includes an overview of the data, results, and discussion; @sec-acknowledgements acknowledges a valuable source of knowledge in creating this report; and @sec-appendix includes supplementary information.
 
@@
 #| tbl-cap: "Outbreaks by type of healthcare location in Toronto in 2023"
-#| tbl-colwidths: [50, 20, 20]
 
@@
 
-@tbl-outbreaks-by-location-type and @fig-outbreaks-by-location-type illustrate that long-term care homes comprise a majority (approximately 56\%) of the total number of healthcare outbreaks in Toronto.
-This is followed by retirement homes (approximately 18\%), chronic care settings in hospitals (approximately 12\%), acute care settings in hospitals (approximately 10\%), psychiatric care settings in hospitals (approximately 2\%), and transitional care locations (approximately 1\%).
+@tbl-outbreaks-by-location-type and @fig-outbreaks-by-location-type illustrate that long-term care homes comprised a majority (approximately 56\%) of the total number of 2023 healthcare outbreaks in Toronto.
+This was followed by retirement homes (approximately 18\%), chronic care settings in hospitals (approximately 12\%), acute care settings in hospitals (approximately 10\%), psychiatric care settings in hospitals (approximately 2\%), and transitional care locations (approximately 1\%).
 Note that these percentages do not add up to 100\% as the proportions in @tbl-outbreaks-by-location-type were rounded to two decimal places.
@@
 #| tbl-cap: "Outbreaks by type of outbreak for Toronto healthcare locations in 2023"
-#| tbl-colwidths: [50, 20, 20]
 
@@
 ```
+
+\newpage
 
@@
 
-@tbl-outbreaks-by-outbreak-type and @fig-outbreaks-by-outbreak-type show that the vast majority of healthcare outbreaks in Toronto are respiratory in nature (approximately 95\%).
-This is followed by enteric outbreaks (approximately 4\%) and other types of outbreaks (less than 1\%).
+@tbl-outbreaks-by-outbreak-type and @fig-outbreaks-by-outbreak-type show that the vast majority of 2023 healthcare outbreaks in Toronto were respiratory in nature (approximately 95\%).
+This was followed by enteric outbreaks (approximately 4\%) and other types of outbreaks (less than 1\%).
 Note that these percentages do not add up to 100\% as the proportions in @tbl-outbreaks-by-outbreak-type were rounded to two decimal places.
+
+\newpage
 
@@
 #| tbl-cap: "Outbreaks by first known cause for Toronto healthcare locations in 2023"
-#| tbl-colwidths: [50, 20, 20]
 
@@
 
-@tbl-outbreaks-by-first-known-cause and @fig-outbreaks-by-first-known-cause illustrate that the majority of Toronto healthcare outbreaks have COVID-19 as their first known cause (approximately 66\%).
-This is followed by rhinovirus (approximately 7\%), respiratory syncytial virus (approximately 5\%), parainfluenza (approximately 3\%), influenza A (approximately 3\%), seasonal coronavirus (approximately 3\%), metapneumovirus (approximately 2\%), norovirus (approximately 2\%), enterovirus/rhinovirus (approximately 1\%), enterovirus (approximately 1\%), CPE (less than 1\%), neonatal Group B streptococcal disease (less than 1\%), and streptococcus pyogenes (less than 1\%).
+@tbl-outbreaks-by-first-known-cause and @fig-outbreaks-by-first-known-cause illustrate that the majority of 2023 Toronto healthcare outbreaks had COVID-19 as their first known cause (approximately 66\%).
+This was followed by rhinovirus (approximately 7\%), respiratory syncytial virus (approximately 5\%), parainfluenza (approximately 3\%), influenza A (approximately 3\%), seasonal coronavirus (approximately 3\%), metapneumovirus (approximately 2\%), norovirus (approximately 2\%), enterovirus/rhinovirus (approximately 1\%), enterovirus (approximately 1\%), CPE (less than 1\%), neonatal Group B streptococcal disease (less than 1\%), and streptococcus pyogenes (less than 1\%).
 Approximately 7\% of outbreaks had a pending or unknown first known cause.
@@
 
-# Acknowledgements {#sec-acknowledgements}
+## Discussion {#sec-data-discussion}
 
-Some of the R code used to create this work was adapted from @tellingstorieswithdata.
-
-# Introduction
-
-You can and should cross-reference sections and sub-sections.
-
-The remainder of this paper is structured as follows. @sec-data....
-
-## Discussion {#data-discussion}
-
-In @sec-data-results, it was observed that long-term care homes had the greatest proportion of total outbreaks among healthcare facilities in Toronto in 2023 (@tbl-outbreaks-by-location-type and @fig-outbreaks-by-location-type), that approximately 95\% of outbreaks were respiratory in nature (@tbl-outbreaks-by-outbreak-type and @fig-outbreaks-by-outbreak-type), and that COVID-19 was the first known cause of a majority of outbreaks (@tbl-outbreaks-by-first-known-cause and @fig-outbreaks-by-first-known-cause).
+In @sec-data-results, it was observed that long-term care homes had the greatest proportion of the total outbreaks among healthcare facilities in Toronto in 2023 (@tbl-outbreaks-by-location-type and @fig-outbreaks-by-location-type), that approximately 95\% of the outbreaks were respiratory in nature (@tbl-outbreaks-by-outbreak-type and @fig-outbreaks-by-outbreak-type), and that COVID-19 was the first known cause of a majority of the outbreaks (@tbl-outbreaks-by-first-known-cause and @fig-outbreaks-by-first-known-cause).
 
@@
 Future research should be conducted to assess the state of disease outbreaks in long-term care homes in particular, both in Toronto and across Canada, in order to gain deeper insight into this critical issue.
-The prevalence of disease outbreaks in other settings, such as schools and workplaces, should also be investigated to expand on this analysis.
+The prevalence of disease outbreaks in additional settings, such as schools and workplaces, should also be investigated to expand on this analysis.
+
+# Acknowledgements {#sec-acknowledgements}
+
+Some of the R code used to create this work was adapted from @tellingstorieswithdata.
 

### 2024-01-22 outputs/paper/paper.qmd
Finalized the paper

@@
 date-format: long
-abstract: "In this report, Toronto Public Health data on 2023 outbreaks in healthcare facilities was analyzed.
+abstract: "In this report, 2023 Toronto Public Health data on outbreaks in healthcare facilities was analyzed.
 Long-term care homes comprised a majority of the outbreaks, approximately 95% of the outbreaks were respiratory, and almost two-thirds of the outbreaks had COVID-19 as their first known cause.
@@
 #| warning: false
+#### Preamble ####
+
+# Purpose: Generates a paper titled "Long-term care homes were hit hardest by
+# 2023 disease outbreaks in Toronto healthcare facilities" in PDF format when
+# rendered.
+# Author: Benny Rochwerg
+# Date: January 23, 2024
+# Contact: 4321benny@gmail.com
+# Pre-requisites: Run the files "01-download_data.R" and "02-data_cleaning.R".
+```
+
+```{r}
+#| echo: false
+#| warning: false
 #### Loading Packages ####
@@
 In March 2020, the World Health Organization proclaimed that the spread of COVID-19 constituted a pandemic [@whocovid].
-Shortly thereafter, in May 2020, the COVID-19 mortality rate in Canadian long-term care homes reached approximately 35\% and represented 81\% of Canadian COVID-19 deaths [@cihi].
+Shortly thereafter, in May 2020, the COVID-19 mortality rate in Canadian long-term care homes, where individuals live and receive healthcare on a continuous basis for reasons such as "advancing age, disability or declining health" [@ltchdefinition], reached approximately 35\% and represented 81\% of Canadian COVID-19 deaths [@cihi].
 In general, the risk of disease circulation is greater in healthcare settings due to an increased presence of ill individuals and nearer physical contact between patients and staff [@cdc].
@@
 Consequently, with the recent advent of large volumes of respiratory diseases such as influenza and COVID-19 in Toronto [@covidflu], assessing the prevalence of disease outbreaks in healthcare facilities in Toronto is of paramount importance.
-As a result, this investigation examines the number of outbreaks by location type (long-term care homes, retirement homes, hospitals, and transitional care facilities), outbreak type (respiratory and enteric/gastroenteric, the latter referring to intestinal and/or stomach illness [@enteric; @gastroenteric]), and first known cause (e.g., COVID-19, influenza A, etc.) to fill this knowledge gap.
+According to @tphdataset, an outbreak is "a localized increase (e.g.
+in an institution, or a specific ward or floor within an institution) in the rate of infection or illness, above that which is expected".
+Thus, this investigation examines the number of outbreaks by location type (long-term care homes, retirement homes, hospitals, and transitional care facilities), outbreak type (respiratory and enteric/gastroenteric, the latter referring to intestinal and/or stomach illness [@enteric; @gastroenteric]), and first known cause (e.g., COVID-19, influenza A, etc.) to fill this knowledge gap.
 
-To do this, Toronto Public Health data on 2023 outbreaks in healthcare facilities was first obtained as described in @sec-data-overview.
+To do this, 2023 Toronto Public Health data on outbreaks in healthcare facilities was first obtained as described in @sec-data-overview.
 Based on this dataset, it was discovered that long-term care homes comprised a majority of the outbreaks, that approximately 95\% of the outbreaks were respiratory, and that almost two-thirds of the outbreaks had COVID-19 as their first known cause (@sec-data-results).
@@
 In doing so, R code was adapted from @tellingstorieswithdata.
+
+\newpage
 
@@
 
+\newpage
+
 ```{r}
@@
 ```
-
-\newpage
 
@@
 Approximately 7\% of outbreaks had a pending or unknown first known cause.
-Note that these percentages do not add up to 100\% as the proportions in @tbl-outbreaks-by-first-known-cause were rounded to two decimal places.
 
@@
 
-Sketches depicting both the desired dataset and the graphs to be generated in this analysis are available in the GitHub Repository.
+Sketches depicting both the desired dataset and the graphs generated in this analysis are available in the GitHub Repository.
 

### 2024-01-22 outputs/paper/paper.qmd
Added average age of long-term care home residents

@@
 All three of these results share common origins.
-For instance, in October 2022, the Government of Ontario removed a masking requirement for caregivers and visitors in long-term care homes [@cbcltch], whose residents are primarily at or above 65 years of age as reported in 2018 [@clri].
+For instance, in October 2022, the Government of Ontario removed a masking requirement for caregivers and visitors in long-term care homes [@cbcltch], whose residents are primarily at or above 65 years of age with an average age of 83 years as reported in 2018 [@clri].
 Since individuals aged 60 years or older are more likely to experience worse health effects after contracting COVID-19, the Public Health Agency of Canada recommends that they wear a respirator or mask to limit the transmission of this pathogen [@phaccovid].

### 2024-01-22 scripts/02-data_cleaning.R
Adding a Data Cleaning subsection to the Appendix section of the paper

@@ -28,7 +28,7 @@ outbreak_cleaned_data <-
          `Type of Outbreak` = `Type of Outbreak`,
          `Outbreak First Known Cause` = `Causative Agent-1`) |>
   
-  # Renaming the Type of Location entries
+  # Renaming the Type of Location entries for clarity
   mutate(`Type of Location` =
            case_match(`Type of Location`,
                       "LTCH" ~ "Long-Term Care Home",

### 2024-01-25 outputs/paper/paper.qmd
Adding a Data Cleaning subsection to the Appendix section of the paper
Removed a space
Added data labels to the graphs and removed the tables

@@
 toc: true
-# The tbl-colwidths parameter used to adjust the table column widths was
-# obtained from Quarto documentation:
-# https://quarto.org/docs/authoring/tables.html
-tbl-colwidths: [50, 20, 20]
 bibliography: references.bib
@@
 # install.packages("here")
-# install.packages("knitr")
 library(tidyverse)
 library(here)
-library(knitr)
 ```
@@
 Based on this dataset, it was discovered that long-term care homes comprised a majority of the outbreaks, that approximately 95\% of the outbreaks were respiratory, and that almost two-thirds of the outbreaks had COVID-19 as their first known cause (@sec-data-results).
-As discussed in @sec-data-discussion, the high prevalence of outbreaks in long-term care homes has likely been exacerbated by the removal of a universal masking requirement, suggesting that the Government of Ontario should consider reinstating this policy to protect these vulnerable individuals.
-With respect to the structure of this paper, @sec-data includes an overview of the data, results, and discussion; @sec-acknowledgements acknowledges a valuable source of knowledge in creating this report; and @sec-appendix includes supplementary information.
+As discussed in @sec-discussion, the high prevalence of outbreaks in long-term care homes has likely been exacerbated by the removal of a universal masking requirement, suggesting that the Government of Ontario should consider reinstating this policy to protect these vulnerable individuals.
+With respect to the structure of this paper, @sec-data includes an overview of the data and the results; @sec-discussion includes a discussion of the results; and @sec-appendix includes supplementary information.
 
@@
 
-After loading the dataset using the R programming language [@citeR] and the `here` package [@citehere], the `tidyverse` [@citetidyverse] and `knitr` [@citeknitr] packages were used to generate tables and graphs.
+After loading the dataset using the R programming language [@citeR] and the `here` package [@citehere], the `tidyverse` [@citetidyverse] package was used to generate graphs.
 In doing so, R code was adapted from @tellingstorieswithdata.
@@
 #| warning: false
-#| label: tbl-outbreaks-by-location-type
-#| tbl-cap: "Outbreaks by type of healthcare location in Toronto in 2023"
 
-#### First Table ####
+#### First Graph Labels ####
 
@@
 
-outbreak_cleaned_data |>
+# Preparing the necessary data to create the labels
+location_data <- outbreak_cleaned_data |>
   count(`Type of Location`) |>
@@
          `Number of outbreaks` = n) |>
-  mutate(`Proportion of total outbreaks` =
-           round(`Number of outbreaks` / nrow(outbreak_cleaned_data), 2)) |>
-  arrange(desc(`Number of outbreaks`)) |>
-  kable()
+  mutate(`Percentage of total outbreaks` =
+           round(100 * `Number of outbreaks` / nrow(outbreak_cleaned_data), 0)) |>
+  arrange(desc(`Number of outbreaks`))
+
+# Generating the labels using the paste function from the Help pane in RStudio
+location_data_labels <- paste(
+    location_data$`Number of outbreaks`, " (", location_data$`Percentage of total outbreaks`, "%)", sep = "")
+
+# Adding the labels to the location data table
+location_data <- location_data |>
+  mutate(`Label` = location_data_labels)
 ```
@@
   # https://stackoverflow.com/questions/25664007/reorder-bars-in-geom-bar-ggplot2-by-value
-  ggplot(mapping = aes(x = reorder(`Type of Location`, n), y = n)) +
+  ggplot(mapping = aes(x = reorder(location_data$`Type of location`, location_data$`Number of outbreaks`), y = location_data$`Number of outbreaks`)) +
   geom_col() +
-  theme_bw() +
+  theme_classic() +
   labs(x = "Type of location", y = "Number of outbreaks") +
-  coord_flip()
+  coord_flip() +
+  # The geom_text function used to add the labels to the graph was obtained from
+  # R Graphics Cookbook:
+  # https://r-graphics.org/recipe-bar-graph-labels
+  # The hjust argument used to align the labels was obtained from
+  # Cédric Scherer's website:
+  # https://www.cedricscherer.com/2021/07/05/a-quick-how-to-on-labelling-bar-graphs-in-ggplot2/
+  geom_text(aes(label = location_data$`Label`), hjust = c(1, 0, 0, 0, 0, 0), colour = "magenta")
 ```
 
-@tbl-outbreaks-by-location-type and @fig-outbreaks-by-location-type illustrate that long-term care homes comprised a majority (approximately 56\%) of the total number of 2023 healthcare outbreaks in Toronto.
+@fig-outbreaks-by-location-type illustrates that long-term care homes comprised a majority (approximately 56\%) of the total number of 2023 healthcare outbreaks in Toronto.
 This was followed by retirement homes (approximately 18\%), chronic care settings in hospitals (approximately 12\%), acute care settings in hospitals (approximately 10\%), psychiatric care settings in hospitals (approximately 2\%), and transitional care locations (approximately 1\%).
-Note that these percentages do not add up to 100\% as the proportions in @tbl-outbreaks-by-location-type were rounded to two decimal places.
+Note that these percentages do not add up to 100\% as they were rounded to the nearest integer.
 
@@
 #| warning: false
-#| label: tbl-outbreaks-by-outbreak-type
-#| tbl-cap: "Outbreaks by type of outbreak for Toronto healthcare locations in 2023"
 
-#### Second Table ####
+#### Second Graph Labels ####
 
@@
 
-outbreak_cleaned_data |>
+# Preparing the necessary data to create the labels
+outbreak_data <- outbreak_cleaned_data |>
   count(`Type of Outbreak`) |>
@@
          `Number of outbreaks` = n) |>
-  mutate(`Proportion of total outbreaks` =
-           round(`Number of outbreaks` / nrow(outbreak_cleaned_data), 2)) |>
-  arrange(desc(`Number of outbreaks`)) |>
-  kable()
+  mutate(`Percentage of total outbreaks` =
+           round(100 * `Number of outbreaks` / nrow(outbreak_cleaned_data), 0)) |>
+  arrange(desc(`Number of outbreaks`))
+
+# Generating the labels using the paste function from the Help pane in RStudio
+outbreak_data_labels <- paste(
+    outbreak_data$`Number of outbreaks`, " (", outbreak_data$`Percentage of total outbreaks`, "%)", sep = "")
+
+# Adding the labels to the outbreak data table
+outbreak_data <- outbreak_data |>
+  mutate(`Label` = outbreak_data_labels)
 ```
@@
   # https://stackoverflow.com/questions/25664007/reorder-bars-in-geom-bar-ggplot2-by-value
-  ggplot(mapping = aes(x = reorder(`Type of Outbreak`, -n), y = n)) +
+  ggplot(mapping = aes(x = reorder(outbreak_data$`Type of outbreak`, -outbreak_data$`Number of outbreaks`), y = outbreak_data$`Number of outbreaks`)) +
   geom_col() +
-  theme_bw() +
-  labs(x = "Type of outbreak", y = "Number of outbreaks")
+  theme_classic() +
+  labs(x = "Type of outbreak", y = "Number of outbreaks") +
+  # The geom_text function and vjust argument used to add the labels to the
+  # graph were obtained from R Graphics Cookbook:
+  # https://r-graphics.org/recipe-bar-graph-labels
+  geom_text(aes(label = outbreak_data$`Label`), vjust = -0.15, colour = "magenta")
 ```
 
-@tbl-outbreaks-by-outbreak-type and @fig-outbreaks-by-outbreak-type show that the vast majority of 2023 healthcare outbreaks in Toronto were respiratory in nature (approximately 95\%).
+@fig-outbreaks-by-outbreak-type shows that the vast majority of 2023 healthcare outbreaks in Toronto were respiratory in nature (approximately 95\%).
 This was followed by enteric outbreaks (approximately 4\%) and other types of outbreaks (less than 1\%).
-Note that these percentages do not add up to 100\% as the proportions in @tbl-outbreaks-by-outbreak-type were rounded to two decimal places.
+Note that these percentages do not add up to 100\% as they were rounded to the nearest integer.
 
@@
 #| warning: false
-#| label: tbl-outbreaks-by-first-known-cause
-#| tbl-cap: "Outbreaks by first known cause for Toronto healthcare locations in 2023"
 
-#### Third Table ####
+#### Third Graph Labels ####
 
@@
 
-outbreak_cleaned_data |>
+# Preparing the necessary data to create the labels
+first_known_cause_data <- outbreak_cleaned_data |>
   count(`Outbreak First Known Cause`) |>
@@
          `Number of outbreaks` = n) |>
-  mutate(`Proportion of total outbreaks` =
-           round(`Number of outbreaks` / nrow(outbreak_cleaned_data), 2)) |>
-  arrange(desc(`Number of outbreaks`)) |>
-  kable()
+  mutate(`Percentage of total outbreaks` =
+           round(100 * `Number of outbreaks` / nrow(outbreak_cleaned_data), 0)) |>
+  arrange(desc(`Number of outbreaks`))
+
+# Generating the labels using the paste function from the Help pane in RStudio
+first_known_cause_data_labels <- paste(" ",
+    first_known_cause_data$`Number of outbreaks`, " (", first_known_cause_data$`Percentage of total outbreaks`, "%)", sep = "")
+
+# Adding the labels to the first known cause data table
+first_known_cause_data <- first_known_cause_data |>
+  mutate(`Label` = first_known_cause_data_labels)
 ```
@@
   # https://stackoverflow.com/questions/25664007/reorder-bars-in-geom-bar-ggplot2-by-value
-  ggplot(mapping = aes(x = reorder(`Outbreak First Known Cause`, n), y = n)) +
+  ggplot(mapping = aes(x = reorder(first_known_cause_data$`Outbreak first known cause`, first_known_cause_data$`Number of outbreaks`), y = first_known_cause_data$`Number of outbreaks`)) +
   geom_col() +
-  theme_bw() +
+  theme_classic() +
   labs(x = "Outbreak first known cause", y = "Number of outbreaks") +
-  coord_flip()
+  coord_flip() +
+  # The geom_text function used to add the labels to the graph was obtained from
+  # R Graphics Cookbook:
+  # https://r-graphics.org/recipe-bar-graph-labels
+  # The hjust argument used to align the labels was obtained from
+  # Cédric Scherer's website:
+  # https://www.cedricscherer.com/2021/07/05/a-quick-how-to-on-labelling-bar-graphs-in-ggplot2/
+  geom_text(aes(label = first_known_cause_data$`Label`), hjust = c(1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), colour = "magenta", size = 3)
 ```
 
-@tbl-outbreaks-by-first-known-cause and @fig-outbreaks-by-first-known-cause illustrate that the majority of 2023 Toronto healthcare outbreaks had COVID-19 as their first known cause (approximately 66\%).
-This was followed by rhinovirus (approximately 7\%), respiratory syncytial virus (approximately 5\%), parainfluenza (approximately 3\%), influenza A (approximately 3\%), seasonal coronavirus (approximately 3\%), metapneumovirus (approximately 2\%), norovirus (approximately 2\%), enterovirus/rhinovirus (approximately 1\%), enterovirus (approximately 1\%), CPE (less than 1\%), neonatal Group B streptococcal disease (less than 1\%), and streptococcus pyogenes (less than 1\%).
+@fig-outbreaks-by-first-known-cause illustrates that the majority of 2023 Toronto healthcare outbreaks had COVID-19 as their first known cause (approximately 66\%).
+This was followed by rhinovirus (approximately 7\%), respiratory syncytial virus (approximately 5\%), parainfluenza (approximately 3\%), influenza A (approximately 3\%), seasonal coronavirus (approximately 3\%), metapneumovirus (approximately 2\%), norovirus (approximately 2\%), enterovirus/rhinovirus (approximately 1\%), enterovirus (approximately 1\%), CPE (less than 1\%), streptococcus pyogenes (less than 1\%), and neonatal Group B streptococcal disease (less than 1\%).
 Approximately 7\% of outbreaks had a pending or unknown first known cause.
 
-## Discussion {#sec-data-discussion}
+# Discussion {#sec-discussion}
 
-In @sec-data-results, it was observed that long-term care homes had the greatest proportion of the total outbreaks among healthcare facilities in Toronto in 2023 (@tbl-outbreaks-by-location-type and @fig-outbreaks-by-location-type), that approximately 95\% of the outbreaks were respiratory in nature (@tbl-outbreaks-by-outbreak-type and @fig-outbreaks-by-outbreak-type), and that COVID-19 was the first known cause of a majority of the outbreaks (@tbl-outbreaks-by-first-known-cause and @fig-outbreaks-by-first-known-cause).
+In @sec-data-results, it was observed that long-term care homes had the greatest proportion of the total outbreaks among healthcare facilities in Toronto in 2023 (@fig-outbreaks-by-location-type), that approximately 95\% of the outbreaks were respiratory in nature (@fig-outbreaks-by-outbreak-type), and that COVID-19 was the first known cause of a majority of the outbreaks (@fig-outbreaks-by-first-known-cause).
 
@@
 This was potentially made worse by the fact that not all of these facilities enforced mask-wearing and that long-term care homes were passed over for personal protective equipment in favour of hospitals [@cbcinquiry].
-Moreover, the fact that several hospitals in Toronto curtailed their mask requirements in June and July of 2023 [@hospitalmasks] likely contributed to the large volume of respiratory disease outbreaks observed in @tbl-outbreaks-by-outbreak-type and @fig-outbreaks-by-outbreak-type.
+Moreover, the fact that several hospitals in Toronto curtailed their mask requirements in June and July of 2023 [@hospitalmasks] likely contributed to the large volume of respiratory disease outbreaks observed in @fig-outbreaks-by-outbreak-type.
 
@@
 The prevalence of disease outbreaks in additional settings, such as schools and workplaces, should also be investigated to expand on this analysis.
-
-# Acknowledgements {#sec-acknowledgements}
-
-Some of the R code used to create this work was adapted from @tellingstorieswithdata.
 
@@
 
+## Data Cleaning {#sec-appendix-cleaning}
+
+The data cleaning process involved filtering out some of the columns from the raw dataset and renaming some of the data entries for clarity and simplicity.
+
 ## Attribution Statement {#sec-appendix-attribution}

### 2024-01-25 scripts/00-simulate_data.R

@@ -10,6 +10,8 @@
 
 #### Loading Packages ####
 
+# install.packages("janitor")
+# install.packages("tidyverse")
 library(janitor)
 library(tidyverse)
 

### 2024-01-25 scripts/01-download_data.R

@@ -9,6 +9,8 @@
 
 #### Loading Packages ####
 
+# install.packages("opendatatoronto")
+# install.packages("tidyverse")
 library(opendatatoronto)
 library(tidyverse)
 

### 2024-01-25 scripts/02-data_cleaning.R

@@ -8,6 +8,7 @@
 
 #### Loading Packages ####
 
+# install.packages("tidyverse")
 library(tidyverse)
 
 #### Cleaning the Dataset ####

### 2024-01-25 scripts/03-test_data.R

@@ -8,6 +8,7 @@
 
 #### Loading Packages ####
 
+# install.packages("tidyverse")
 library(tidyverse)
 
 #### Testing the Cleaned Dataset ####
